The four postulates of quantum mechanics are three

### 2020-11-25 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Another pass to the tensor article

	\begin{defn}[Composite systems]\label{def_comp}
		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
		\begin{enumerate}
			\item Every preparation of both subsystems is a preparation of the
			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
			for C, there exists a map (not yet specified)
			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states. In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
			\item For every preparation of the composite, local projective measurements must have at least one
			outcome with non-zero probability. Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$. 
		\end{enumerate}
		It is important to understand that these requirements are necessary. Requirement 1 ensures that the composite system is well defined at least when the components are prepared independently. Conceptually, this ensures that the composite system contains all the properties of the components. Note that superselection rules or other restrictions may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state). The tensor product is recovered only when all pairs are compatible. Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.
	\end{defn}

>>>>

	\begin{defn}[Composite systems]\label{def_comp}
		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
		\begin{enumerate}
			\item Every preparation of both subsystems is a preparation of the
			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
			for C, there exists a map (not yet specified)
			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states. In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
			\item For every preparation of the composite, local projective measurements must have at least one
			outcome with non-zero probability. Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$. 
		\end{enumerate}
		It is important to understand that these requirements are necessary. Requirement 1 ensures that the composite system is well defined at least when the components are prepared independently. Conceptually, this ensures that the composite system contains all the properties of the components. Note that superselection rules or other restrictions may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state). The tensor product is recovered only when all pairs are compatible. Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components. Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems. That is, while the systems $A$ and $B$ by itself would define an observable $O_A$ and $O_B$, when put grouped together all values would be assigned zero probability; those observables no longer exist. In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
	\end{defn}

### 2020-11-25 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Another pass to the tensor article

	We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}\comment{da mettere altrove:, namely, 
		its state is fully described by the probability amplitudes of the system's observables and nothing else}. The first key insight is that the first two postulates of quantum theory (introduced below) already assume that, at least in some cases, the probability distribution of the
	measurement outcomes of one system is independent of the other systems' (statistical independence). We cannot even talk about a system in the first place if we cannot characterize it independently. The second key insight is that, since all measurement
	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case. These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states). So the work consists of two interrelated efforts: a physical argument that starts from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.

@@

	TODO: pensarci con calma: riscrivere il paragrafo seguente, magari partendo da un esempio e introducendo il concetto seguente:
	Frase in cui diciamo che abbiamo bisogno di projective spaces perche' i vettori $|\psi\>$ non rappresentano proprieta' (direbbe Lorenzo) o events (o proposizioni) in probability space (direbbe Gabriele) ma sono i sottospazi che lo fanno.

	Since quantum states are rays in Hilbert space, we need projective spaces
	Events in probability space do not correspond to vectors in the Hilbert space, but to subspaces, which means we will need elements of projective geometry. If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$. Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }. In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$. The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projective spaces. The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.

@@

	\togli{This axiomatization implicitly contains a definition of
		``quantum system'' which is crucial for what follows, so we need to
		clarify the assumptions that it contains. We will use the following
		definition for a quantum
		system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
			of freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of
			dimension $d$ which contains all the states that describe the
			values of its possible properties. In accordance with the
			postulate (a), these values correspond to a basis of the space,
			given by the eigenvectors of the observable corresponding to that
			property''}. \commentg{We may have to revise to be more clear.
			Where is this used?} } As discussed, we limit ourselves to
	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
	state, {\em unconditioned on anything else}. In particular, we note
	that restrictions due to superselection rules arise either from
	practical (not fundamental) limitations on the actions of the
	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
	ill-defined quantum systems. \footnote{For example, in situations where
		indistinguishability plays a role (e.g.~in QFT), one cannot consider
		an electron as a quantum system: in this case the quantum system is
		the field. The electron is an excitation of the field, {\em not} a
		(sub)system.} We call this condition ``preparation independence''.  [We
	emphasize that the kinematic independence is inequivalent to dynamical
	independence (or isolation).  Indeed if two systems interact, their
	interaction may lead to dynamical restrictions in the state spaces.
	Here we will not consider dynamical evolution, which is contained in
	postulate (d).]

@@

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Two events $a$ and $b$ are defined to be 
	independent given $z$ if $p(a\wedge b|z)=p(a|z)\:p(b|z)$, which is true if and only if $p(b|z\wedge a)=p(b|z)$. If $a$ and $b$ are measurement outcomes, the probability of $b$ is not a function of $a$.

	Riscrivere questo: scritto cosi', sembra che la regola di Born non si possa applicare sempre....
	We note that the Born
	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$). Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$. In what cases are we allowed to use the Born rule? That is, in what cases can we say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$? Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$, where the systems  are statistically independent. In this sense, postulate (b) already implies  ``statistical independence'' between
	different systems if they are prepared independently. Namely, the
	outcome probability of one is independent of any properties of any
	other in the absence of any dynamical coupling and supposing
	independent preparation. [Of course, the dynamics may couple the state
	of the system to other systems so that the outcomes may depend on what
	happens to other systems.  Analogously, one may prepare the system in
	a way that depends on other systems. But we consider independent
	preparations and do not consider dynamics, as it is sufficient for our
	aims.] We can then formalize ``statistical independence'' of
	independently prepared systems by saying that their Born rule
	satisfies:
	\begin{eqnarray}
	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
		A})\:p(b|\psi_{\cal B})
	\labell{bindip}\;,
	\end{eqnarray}
	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
	B$ respectively. The first equality in equation \eqref{bindip} embodies the
	definition of statistical independence, the second follows from the
	form of the Born rule for independently prepared systems (the
	probability depends only on the state and observables of each system
	on its own).

	A particular case is when one system is
	prepared in a fixed eigenstate $|b\>$ of the observable measured on it,
	$p(a\wedge b|\psi\wedge b)=p(a|\psi)\:p(b|b)=p(a|\psi)$, where the last equality follows from the Born rule since $\<b|b\>=1$. 
	If we define $M_b(\pj{a}) = M(\pj{a},\pj{b})$, substituting the values of the
	probabilities from the Born rule, we have:
	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
	=\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
	\nonumber \\&
	=\left|\<a|\psi\>_{\cal A}\right|^2
	\labell{questa},
	\end{align}
	where the first and second terms contain the inner product in the composite
	space $\cal C$. [This is not a new assumption: it follows from the
	measurement postulate (b) for the composite system.] This means that,
	when one subsystem is prepared in an eigenstate of what is measured
	there, the state space of the other is mapped preserving the square of
	the inner product.
	This implies orthogonality and
	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
	colinear transformation by definition. Geometrically, recall that $M_b$ maps rays to rays. The fact that $M_b$ is colinear means that it also maps higher order subspaces to higher order subspaces (lines to lines, planes to planes, and so on) while preserving inclusion (if a line is within a plane, the mapped line will be within the mapped plane). In this case, the fundamental
	theorem of projective geometry \cite{fun} applies, which tells us that
	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
	Moreover, conservation of probability further constrains it to be
	either linear or antilinear. This tells us that the corresponding $m$
	is either linear or antilinear in the first argument. Namely, if equation
	\eqref{questa} holds, then
	\begin{align}
	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
	\\\mbox{ or }
	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
	\end{align}
	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored. Given a Hilbert space, in fact, we can imagine replacing all vectors and all the operators with their Hermitian conjugate, mapping vectors into duals $|\psi\>^\dag=\<\psi|$. These changes would effectively cancel out leaving the physics unchanged. The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$.  (For example, in his first papers
	Schr\"odinger used both signs in his equation: effectively writing
	{\em two} equivalent equations with complex-conjugate solutions
	\cite{sch}). We can repeat the same analysis for the second
	argument of $m$ to conclude that it is a bilinear map, condition (H2).

>>>>

	We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems. The composite system is therefore
	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}.\footnote{Note that this does not exclude entangled states since they too are characterized by the observables and only the observables of the constituents.} The first key insight is that the first two postulates of quantum theory (introduced below) already assume that the preparation of one system is independent from the preparation of another (statistical independence). In fact, we cannot even talk about a system in the first place if we cannot characterize it independently. The second key insight is that, since all measurement
	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case. These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states). So the work consists of two interrelated efforts: a physical argument that starts from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.

@@

	To make these arguments precise, we will need to use elements of projective geometry.
	If $\psi$ is a vector in our Hilbert space, the pure states is represented by the ray $\pj{\psi}$, which is the subspace spanned by $\psi$. This includes not only vectors with arbitrary phase, but also arbitrary modulus. The term ``ray'' can therefore be misleading since $\pj{\psi}$ is really a complex { \em plane }. In the same way, constraining the observable $X$ to a particular value $x_0$ means identifying the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$. As we need to map states and observables between subsystems and composite system, the map $M$ derived from the probability space is really a map between subspaces, between the projective spaces. The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.

@@

	\togli{This axiomatization implicitly contains a definition of
		``quantum system'' which is crucial for what follows, so we need to
		clarify the assumptions that it contains. We will use the following
		definition for a quantum
		system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
			of freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of
			dimension $d$ which contains all the states that describe the
			values of its possible properties. In accordance with the
			postulate (a), these values correspond to a basis of the space,
			given by the eigenvectors of the observable corresponding to that
			property''}. \commentg{We may have to revise to be more clear.
			Where is this used?} }

	Note that we limit ourselves to
	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
	state, {\em unconditioned on anything else}. We call this condition ``preparation independence'' and it should be noted that the tensor product applies only in this case. For example, the composite system of two electrons is not the tensor product, rather the anti-symmetrized tensor product, precisely because the second electron cannot be prepared in the same state of the first. We note
	that restrictions due to superselection rules arise either from
	practical (not fundamental) limitations on the actions of the
	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
	ill-defined quantum systems. In the example above, the field is the proper quantum system and the electrons are its excitations. \footnote{We
	emphasize that the kinematic independence is inequivalent to dynamical
	independence (or isolation).  Indeed if two systems interact, their
	interaction may lead to dynamical restrictions in the state spaces.
	Here we will not consider dynamical evolution, which is contained in
	postulate (d).}

@@

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Consider $p(a \wedge b | \psi \wedge b)$ which represents the probability of measuring $a$ on system $A$ and $b$ on system $B$ having prepared system $A$ in $\psi$ and system $B$ in $b$. We have $p(a \wedge b | \psi \wedge b) = p(a | \psi \wedge b \wedge b) p(b | \psi \wedge b) = p(a | \psi \wedge b ) p(b | \psi \wedge b)$. The Born rule tells us that $p(a | \psi \wedge b) = |\<a|\psi\>|^2$ and that $p(b | \psi \wedge b) = |\<b|b\>|^2 = 1$, where $a$, $b$ and $\psi$ are assumed to be normalized. We have:
	\begin{align}
	p(a\wedge b|\psi \wedge b)&=p(a|\psi)\\
	p(a\wedge b|a \wedge \phi)&=p(b|\phi)
	\end{align}
	In other words, since the probability for a measurement on one system depends only on its pure state, the Born rule automatically assume that the measurement of one system is independent from the preparation of the other. We call this property ``statistical independence''.\footnote{One can also prove that the measurements on the components are independent as well, but we only strictly need preparation.}

	We can use this fact to characterize our map $M$. Note that $M(\pj{a}, \pj{b})$ will correspond to the composite state where $A$ and $B$ are prepared in the respective states. We can define define $M_b(\pj{a}) = M(\pj{a},\pj{b})$ and, substituting the values of the
	probabilities from the Born rule, we have:
	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
	=\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
	\nonumber \\&
	=\left|\<a|\psi\>_{\cal A}\right|^2
	\labell{questa},
	\end{align}
	where the first and second terms contain the inner product in the composite
	space $\cal C$. [This is not a new assumption: it follows from the
	measurement postulate (b) for the composite system.] This means that,
	when one subsystem is prepared in an eigenstate of what is measured
	there, the state space of the other is mapped preserving the square of
	the inner product.
	This implies orthogonality and
	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
	colinear transformation by definition. Geometrically, recall that $M_b$ maps rays to rays. The fact that $M_b$ is colinear means that it also maps higher order subspaces to higher order subspaces (lines to lines, planes to planes, and so on) while preserving inclusion (if a line is within a plane, the mapped line will be within the mapped plane). In this case, the fundamental
	theorem of projective geometry \cite{fun} applies, which tells us that
	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
	Moreover, conservation of probability further constrains it to be
	either linear or antilinear. This tells us that the corresponding $m$
	is either linear or antilinear in the first argument. Namely, if equation
	\eqref{questa} holds, then
	\begin{align}
	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
	\\\mbox{ or }
	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
	\end{align}
	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored. Given a Hilbert space, in fact, we can imagine replacing all vectors and all the operators with their Hermitian conjugate, mapping vectors into duals $|\psi\>^\dag=\<\psi|$. These changes would effectively cancel out leaving the physics unchanged. The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$ are equivalent.  (For example, in his first papers
	Schr\"odinger used both signs in his equation: effectively writing
	{\em two} equivalent equations with complex-conjugate solutions
	\cite{sch}). We can repeat the same analysis for the second
	argument of $m$ to conclude that it is a bilinear map, condition (H2).
